---
title: "Script 4: Data Analysis and Modeling"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}

  knitr::opts_chunk$set(echo = TRUE, eval=FALSE)

```

### Introduction

This document describes all of the code used to the generate the output included in the 'To Airbnb?' study. This analysis uses the output from the **3_data_prep.rmd** file found in the same online repository. 

## Preliminary Commands

&nbsp;

We begin with a number of preliminary commands.  First we load all of the necessary R libraries for this analysis. 

```{r load_libraries, message=FALSE, warning=FALSE, comment=FALSE, echo=TRUE, results='hide'}

  library(tidyverse)
  library(sf)
  library(geosphere)
  library(ggmap)
  library(multidplyr)
  library(tmap)

```

```{r set_options, message=FALSE, warning=FALSE, comment=FALSE, echo=TRUE, results='hide'}

  options(stringsAsFactors=FALSE)
  options(xtable.comment = FALSE)
  options(xtable.caption.placement='top')

```

Next, we source the necessary custom code files for the analysis.  

```{r source_files, message=FALSE, warning=FALSE, comment=FALSE, echo=TRUE, results='hide'}

  source(file.path(getwd(), 'functions', 'abb_Functions.R'))

```

### Set Parameters

```{r set_constants, message=FALSE, warning=FALSE, comment=FALSE, echo=TRUE, results='hide'}
  
  # Exchange rate: AUS to US 
  exch_rate <- 1.32  

  # Date of investment decision
  decision_date <- lubridate::as_date('2015-09-01')
  
```

## Load Data

We then load the data.

```{r load_data, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE} 

  load(file.path(getwd(), 'data', 'prepared', 'prepared_data.RData'))
  str_col = col_df$rgb[1]
  
```

#### Compare Revenue Distributions

```{r flatten_ltr}

  ltr_df <- ltr_tdf %>% 
    unnest() %>%
    select(-suburb.x) %>%
    rename(suburb=suburb.y)

```

Before we examine the distributions of the observed short-term revenues and the estimated long-term revenues, we must produce the estimates of long-term revenue for the short-term properties.

We start by specifying the imputation model specification we will use to estimated likely long-term (weekly) rental prices for each short-term property.

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  ltr_mod_spec <- formula(log(price) ~ as.factor(type) +
                            as.factor(bedbath) + as.factor(suburb) +
                            as.factor(month))

```

We then employ a custom function to impute the long term rents for the short-term properties.  The 'clip.field' is a spatial fixed effects that add spatial recognition to the models.

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  str_tdf$tbg <- paste0(str_tdf$product,'_', str_tdf$geo_mrkt)
  ltr_df$tbg <- paste0(ltr_df$product,'_', ltr_df$geo_mrkt)

  imp_df <- abbImputeCompare(str.df=str_tdf, 
                                ltr.df=ltr_df, 
                                mod.spec=ltr_mod_spec, 
                                match.factor='suburb',
                                split.field='tbg', 
                                verbose=FALSE) 
  
```

Next, we also impute prices via a KNN approach

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  kimp_df <- knnImputeWrap(ltr_df = ltr_df, 
                           str_df = str_tdf,
                           knns = c(7,7))
  
  imp_df$ltr_imp_revenueK <- kimp_df$year[match(imp_df$property_id,
                                               kimp_df$property_id)]

```

We then merge the imputed values

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  imp_df <- imp_df %>%
    rename(ltr_imp_revenueH = ltr_imp_revenue) %>%
    mutate(ltr_imp_revenueK = ((365-imp_dom)/365) * ltr_imp_revenueK,
           ltr_imp_revenue = .5 * (ltr_imp_revenueH + ltr_imp_revenueK))
  
  imp_df <- imp_df %>%
    left_join(str_tdf %>%
                dplyr::select(property_id, bedrooms, bathrooms, type, latitude,
                              longitude),
              by='property_id')
 
```

From the imputed results, we extract the imputed rental values and merge to the original short-term data.

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  # Extract out DFs (only those that imputed)
  str_tdf <- str_tdf %>%
    dplyr::left_join(imp_df %>%
                       dplyr::select(-c(bedrooms, bathrooms, type, latitude, longitude)),
                     by='property_id') %>%
    dplyr::mutate(act_pref = ifelse(revenue_net > ltr_imp_revenue, 1, 0),
                  # lik_pref = ifelse(lik_revenue_net > ltr_imp_revenue, 1, 0),
                  pot_pref = ifelse(pot_revenue_net > ltr_imp_revenue, 1, 0),
                  pot_pref_mc = ifelse(pot_revenue_net * .85 > ltr_imp_revenue, 1, 0),
                  pot_pref_us = ifelse(pot_revenue_net > ltr_imp_revenue * .85, 1, 0))

```

## Logit models

Now we attempt to explain the drivers of preference with a set of logistic regression models.  

First, a basic model with just structural characteristics. We estimate using the standard dependent variable as well as the two sensitivity scenarios. 

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  glm_pot_str <- glm(pot_pref~type+as.factor(bedbath),
                       family=binomial(link='logit'),
                       data=str_tdf)
  glm_pot_str_mc <- glm(pot_pref_mc~type+as.factor(bedbath),
                       family=binomial(link='logit'),
                       data=str_tdf)
  glm_pot_str_us <- glm(pot_pref_us~type+as.factor(bedbath),
                       family=binomial(link='logit'),
                       data=str_tdf)
  
```

Then a model that adds in the geographic submarkets

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  glm_pot_loc <- glm(pot_pref ~ type + as.factor(bedbath)+ as.factor(geo_mrkt),
                     family=binomial(link='logit'),
                     data=str_tdf)
  glm_pot_loc_mc <- glm(pot_pref_mc ~ type + as.factor(bedbath)+ as.factor(geo_mrkt),
                     family=binomial(link='logit'),
                     data=str_tdf)
  glm_pot_loc_us <- glm(pot_pref_us ~ type + as.factor(bedbath)+ as.factor(geo_mrkt),
                     family=binomial(link='logit'),
                     data=str_tdf)
```

Finally a model that adds in host policies. 

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}


  glm_pot_host <- glm(pot_pref ~ type + as.factor(bedbath) + as.factor(geo_mrkt) +
                        I(max_guests/bedrooms) + min_stay + I(canc_policy == 'Flexible'),
                      family=binomial(link='logit'),
                      data=str_tdf)
  glm_pot_host_mc <- glm(pot_pref_mc ~ type + as.factor(bedbath) + as.factor(geo_mrkt) +
                        I(max_guests/bedrooms) + min_stay + I(canc_policy == 'Flexible'),
                      family=binomial(link='logit'),
                      data=str_tdf)
  glm_pot_host_us <- glm(pot_pref_us ~ type + as.factor(bedbath) + as.factor(geo_mrkt) +
                        I(max_guests/bedrooms) + min_stay + I(canc_policy == 'Flexible'),
                      family=binomial(link='logit'),
                      data=str_tdf)
  
```

We package these all together into a lists of results for eacy storage

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=TRUE}

  pot_glms <- list(str=glm_pot_str,
                   loc=glm_pot_loc,
                   host=glm_pot_host)
  pot_mc_glms <- list(str=glm_pot_str_mc,
                   loc=glm_pot_loc_mc,
                   host=glm_pot_host_mc)
  pot_us_glms <- list(str=glm_pot_str_us,
                   loc=glm_pot_loc_us,
                   host=glm_pot_host_us)
  

```

### Conditional Expectations

Here we create the data for producing conditional expectation charts.  This involves creating a set of basket or exemplar scoring properties to use as the permutations of conditions.

```{r}

# Create New variables

 str_tdf$gpb <- str_tdf$max_guests / str_tdf$bedrooms
 str_tdf$canc <- ifelse(str_tdf$canc_policy == 'Flexible', 1, 0)

# Recreate Model 
  
 pred_model <- glm(pot_pref ~ type + as.factor(bedbath) + as.factor(geo_mrkt) +
                    gpb + min_stay + canc,
                      family=binomial(link='logit'),
                      data=str_tdf)
  

### Create variables for pred 

  bedbath <- unique(str_tdf$bedbath)
  type <- unique(str_tdf$type)
  geo_mrkt <- unique(str_tdf$geo_mrkt)
  canc <- c(0,1)
  min_stay <- c(1,2,3,4,7)
  gpb <- 1:5
  
  pred_df <- expand.grid(bedbath, type, geo_mrkt, canc, min_stay, gpb)
  names(pred_df) <- c('bedbath', 'type', 'geo_mrkt', 'canc', 'min_stay', 'gpb')
  
  pred_df$id <- apply(pred_df, 1, function(x) paste0(substr(x['type'], 1,1),
                                                     x['bedbath'], 
                                                     substr(x['geo_mrkt'], 1, 5)))
  
  str_tdf$id <- apply(str_tdf, 1, function(x) paste0(substr(x['type'], 1,1),
                                                     x['bedbath'], 
                                                     substr(x['geo_mrkt'], 1, 5)))
  pred_df$aid <- apply(pred_df, 1, function(x) paste0(x['bedbath'], 
                                                      substr(x['geo_mrkt'], 1, 5)))
  str_tdf$aid <- apply(str_tdf, 1, function(x) paste0(x['bedbath'], 
                                                      substr(x['geo_mrkt'], 1, 5)))
  
```

We then make the predictions over all of the exemplar properties

```{r}

# Make predictions
  pred_df$raw <- predict(pred_model, newdata = pred_df)
  pred_df$prob <-  exp(pred_df$raw)/(1+exp(pred_df$raw))
  pred_df$sh <- ifelse(pred_df$prob > .5, 1, 0)
  pred_df$sh <- factor(pred_df$sh, levels = c(1,0))
  
  pred_df$gpb <- as.factor(pred_df$gpb)
  pred_df$cc <- ifelse(pred_df$canc == 1, 'Flexible', 'Strict')
  pred_df$ct <- paste0(ifelse(pred_df$type == 'House', 'House', 'Apt'), 
                       " - ",  pred_df$cc)

  pred_df$ms <- as.character(pred_df$min_stay)
  pred_df$ms[pred_df$ms == '4'] <- '4-6'
  pred_df$ms[pred_df$ms == '7'] <- '7+'
  pred_df$ms <- factor(pred_df$ms, levels=c('1', '2', '3', '4-6', '7+'))
  
  pred_df$sm <- as.character(pred_df$geo_mrkt)
  pred_df$sm[pred_df$sm == 'city-core'] <- 'City-Core'
  pred_df$sm[pred_df$sm == 'city'] <- 'City'
  pred_df$sm[pred_df$sm == 'beach'] <- 'Beach'
  pred_df$sm[pred_df$sm == 'suburban'] <- 'Suburban'
  pred_df$sm <- factor(pred_df$sm, levels = c('City-Core', 'City', 'Beach', 'Suburban'))

  pred_df$ct <- factor(pred_df$ct, levels = c('Apt - Strict', 'House - Strict', 
                                              'Apt - Flexible', 'House - Flexible'))
    
```

We save these outputs for use in the data visualization stage next.  

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=TRUE}

  dir.create(file.path(getwd(), 'data', 'analyzed'))
  save(str_tdf, ltr_tdf, ltr_df, col_df, suburb_sf, studyarea_sf, sa1_sf, str_sf, ltr_sf,
       pot_glms, core_suburbs, str_tdf_all, pot_mc_glms, pot_us_glms,
       pred_df, 
       file=file.path(getwd(), 'data', 'analyzed', 'analyzed_data.RData'))
  
```
