---
title: "Script 5: Data Visualization"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
  
  knitr::opts_chunk$set(echo = TRUE)

```

Load the necessary libraries

```{r set_paths, message=FALSE, warning=FALSE, comment=FALSE, echo=TRUE, results='hide'}

  library(tmap)
  library(magrittr)
  library(dplyr)
  library(ggplot2)

  source(file.path(getwd(), 'functions', "abb_Functions.R"))

```

Next we load the data objects that we built earlier.

```{r load_data, message=FALSE, warning=FALSE, comment=FALSE, cache=TRUE, echo=FALSE, results='hide'}
  load(file.path(getwd(), 'data', 'prepared', 'prepared_data.RData'))

  load(file.path(getwd(), 'data', 'analyzed', 'analyzed_data.RData'))

  dir.create(file.path(getwd(), 'images'))

  str_col = col_df$rgb[1]
    
```

### Host type investigation

We look at host behavior, in aiming to identify cases the profit maximizing hosts. 

```{r host_types}

  str.col <- col_df$rgb[1]

  ## Create table of host block frequencies

  host.block.freq <- ggplot(str_tdf_all, aes(x=block_rate)) +
     geom_density(fill=str.col, color=str.col) +
     ylab('Frequency\n') +
     xlab('\n\n% of Time Property is Blocked') +
     scale_x_continuous(breaks=seq(0,1,.2),
                        labels=c('0%', '20%', '40%', '60%', '80%', '100%')) +
     #ggtitle('Host Block Rate Frequency\n') +
     theme(plot.title = element_text(hjust = 0.5, size = 10)) +
     theme(axis.text.y=element_blank(),
           axis.ticks.y=element_blank(),
           plot.caption=element_text(size=6, hjust=0)) + 
     labs(caption = paste0('\nDensity plot of frequency of blocking schedule for short term ',
                           '\nleasing observations. Blocking indicates that the property ',
                           'is \nnot available for nightly rental on Airbnb. Vertical lines',
                           '\ndelineate the three host type groupings.'),
          title = 'Host Blocking Frequency\n') +
    geom_vline(xintercept = .25, linetype="dashed", color = "black", size=.5, alpha=.5)+
    geom_vline(xintercept = .75, linetype="dashed", color = "black", size=.5, alpha=.5)
    
  
  block.rate.period <- ggplot(str_tdf_all, 
                              aes(x=block_rate, y=nbr_block * (366/total_days))) +
    geom_point(fill=alpha(str.col,.3), color=alpha(str.col,.3), size=.1, shape=16) +
    ylab('Nbr of Blocked Periods\n') +
    xlab('\n\n% of Time Property is Blocked') +
    scale_x_continuous(breaks=seq(0, 1, .2),
                       labels=c('0%', '20%', '40%', '60%', '80%', '100%')) +
    #ggtitle('Block Rate vs. Nbr. of Blocked Periods\n') +
     theme(plot.title = element_text(size = 10, hjust = 0.5),
           axis.text.y=element_blank(),
           axis.ticks.y=element_blank(),
           plot.caption=element_text(size=6, hjust=0)) + 
     labs(caption = paste0('\nCumulative time of blocking for short term properties ',
                           '(x-axis) \nversus the total number of periods where the ',
                           'property was \nblocked from nightly rental.',
                           'Vertical lines delineate the three \nhost type groupings.'),
          title = 'Block Rate vs Blocked Periods\n')+
    geom_vline(xintercept = .25, linetype="dashed", color = "black", size=.5, alpha=.5)+
    geom_vline(xintercept = .75, linetype="dashed", color = "black", size=.5, alpha=.5)
  
    png(filename = file.path(getwd(), 'images', "block_rate.png"), 
       width = 5.5, height = 3, units = 'in', res = 300)

       ggMultiPlots(host.block.freq, block.rate.period, cols=2)
       
    dev.off()  
  
```

```{r host_types2}

  base.host <- ggplot(str_tdf_all,
                    aes(x=block_rate, y=blockpertime)) +
                geom_point(alpha=.7, size=.4, shape=16, color='gray70') +
                ylab('# Blckd Periods\n') +
                xlab('\n% of Time Property is Blocked') +
                ggtitle('Block Rate vs. Nbr. of Blocked Periods\n') +
                scale_x_continuous(breaks=seq(0, 1, .2),
                                   labels=c('0%', '20%', '40%',
                                            '60%', '80%', '100%')) +
                #scale_color_manual(values=c(abb.col[2], abb.col[4],
                #                            abb.col[7], abb.col[9])) +
                theme(legend.position='bottom',
                      legend.title = element_blank(),
                      plot.title = element_text(hjust = 0.5))+
                guides(colour = guide_legend(override.aes = list(size=3,                                                                       alpha=1)))

   ps.host <- base.host +
    geom_point(data=str_tdf_all[str_tdf_all$host_type == 'Profit Seeker', ], size=.4)  +
    ggtitle('Profit Seekers')

  os.host <- base.host +
    geom_point(data=str_tdf_all[str_tdf_all$host_type == 'Opportunistic Sharer', ],
               size=.4)  +
    ggtitle('Opportunistic Sharer')

  un.host <- base.host +
    geom_point(data=str_tdf_all[str_tdf_all$host_type == 'Unknown', ], size=.4)  +
    ggtitle('Unknown')

  png(filename = file.path(getwd(), 'images', "host_scat.png"), 
       width = 4, height = 6.5, units = 'in', res = 300)

    ggMultiPlots(ps.host, un.host, os.host, cols=1)

  dev.off()  
     
```

```{r plot_hostbar, message=FALSE, warning=FALSE, comment=FALSE, echo=FALSE, fig.height=3, fig.width=4, fig.cap='\\label{htc}Host Type Counts'}

 ## Build plot

  png(filename = file.path(getwd(), 'images', "host_bar.png"), 
       width = 3, height = 3, units = 'in', res = 300)

 ggplot(str_tdf_all, aes(host_type, fill=host_type)) +
    geom_bar() +
    scale_fill_manual(values=c(col_df$rgb[2], col_df$rgb[4], col_df$rgb[9]))+
    theme(legend.position='none') +
    ylab('') +
    xlab('') +
     theme(plot.title = element_text(size = 10, hjust = 0.5),
           axis.text.y=element_blank(),
           axis.text.x=element_text(size = 6),
           axis.ticks.y=element_blank(),
           plot.caption=element_text(size=6, hjust=0)) + 
     labs(caption = paste0('\nBreakdown of Host Types for Short Term Properties:\n\n',
                           '1) Profit Seekers:               <= 25% Blocked             3,056 Obs.\n',
                           '2) Opportunistic Sharers:   >= 75% Blocked             1,743 Obs.\n',
                           '3) Unknowns:                     25% > 75% Blocked       2,707 Obs.\n')) + 
   annotate("text", x = 1, y = 3300, label = "41%") +
   annotate("text", x = 2, y = 2000, label = "23%") +
   annotate("text", x = 3, y = 2949, label = "36%") +
   coord_cartesian(ylim=c(0,3500))

  dev.off()  
  
```
### Observation Bar Chart

Here we make bar charts breaking down the number of observations by type and submarket. 

```{r smtype.plot, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=TRUE, fig.width=9}
 
  sm_col <- c(col_df$rgb[3], col_df$rgb[4], col_df$rgb[5], 
              col_df$rgb[6])
  
   # Make Airbnb submarket map
   str_tdf$Submarket <- as.character(str_tdf$geo_mrkt)
   str_tdf$Submarket[str_tdf$Submarket == 'city-core'] <- '   City-Core   '
   str_tdf$Submarket[str_tdf$Submarket == 'city'] <- '   City   '
   str_tdf$Submarket[str_tdf$Submarket == 'beach'] <- '   Beach   '
   str_tdf$Submarket[str_tdf$Submarket == 'suburban'] <- '   Suburban   '
   str_tdf$Submarket <- factor(str_tdf$Submarket, 
                                  levels = c('   City-Core   ', '   City   ', 
                                             '   Suburban   ', '   Beach   '))
   
   strtype_bar <- ggplot() +
     geom_bar(data=str_tdf, aes(x=Submarket, fill=Submarket)) +
     scale_fill_manual(values=sm_col, name='Submarket') +
     facet_wrap(~type) +
     ylab('# of Properties\n') + xlab('')+
     theme(legend.position = 'none',
           legend.title = element_blank(),
           axis.ticks.x=element_blank()) +
     ggtitle('Short Term Properties')+
     theme(plot.title = element_text(size = 10, hjust = 0.5),
           axis.text.y=element_blank(),
           axis.text.x=element_text(size = 6),
           axis.ticks.y=element_blank(),
           plot.caption=element_text(size=6, hjust=.5)) + 
     labs(caption = paste0('\nBreakdown of 2,415 short-term properties by submarket and ',
                           ' structure type.\n'))
  # strtype_bar
   
   # Make Long term rental map
 
   ltr_tdf$Submarket <- as.character(ltr_tdf$geo_mrkt)
   ltr_tdf$Submarket[ltr_tdf$Submarket == 'city-core'] <- '   City-Core   '
   ltr_tdf$Submarket[ltr_tdf$Submarket == 'city'] <- '   City   '
   ltr_tdf$Submarket[ltr_tdf$Submarket == 'beach'] <- '   Beach   '
   ltr_tdf$Submarket[ltr_tdf$Submarket == 'suburban'] <- '   Suburban   '
   ltr_tdf$Submarket <- factor(ltr_tdf$Submarket, 
                                  levels = c('   City-Core   ', '   City   ', 
                                             '   Suburban   ', '   Beach   '))
   ltrtype_bar <- ggplot() +
     geom_bar(data=ltr_tdf, aes(x=Submarket, fill=Submarket)) +
     scale_fill_manual(values=sm_col) +
     facet_wrap(~type) +
     ylab('# of Properties\n') + xlab('')+
     theme(legend.position = 'none',
           legend.title = element_blank(),
           axis.text.x=element_blank(),
           axis.ticks.x=element_blank()) +
     ggtitle('Long Term Properties')+
     theme(plot.title = element_text(size = 10, hjust = 0.5),
           axis.text.y=element_blank(),
           axis.text.x=element_text(size = 6),
           axis.ticks.y=element_blank(),
           plot.caption=element_text(size=6, hjust=.5)) + 
     labs(caption = paste0('\nBreakdown of 64,305 long-term properties by submarket and ',
                           ' structure type\n'))
     #ltrtype_bar
   
     png(filename = file.path(getwd(), 'images', "loc_bar.png"), 
       width = 8, height = 4, units = 'in', res = 300)
 
     ggMultiPlots(strtype_bar, ltrtype_bar, cols=2)
     dev.off()  
 
       
```
## Suburb Location Map

We start by create the maps of the observation locations, the study area and the suburb boundaries. 

First the suburb boundaries and study areas. 

```{r sub_loc.map, message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=TRUE}
 
  # Set submarket colors
  suburb_sf$col <- 'gray80'
  suburb_sf$col[suburb_sf$class == 'Middle'] <- 'gray50'
  suburb_sf$col[suburb_sf$class == 'Inner'] <- 'gray20'
  
  inn_sf <- suburb_sf %>% dplyr::filter(class == 'Inner') %>% st_union()
  mid_sf <- suburb_sf %>% dplyr::filter(class == 'Middle' & 
                                          name != 'Seabrook' &  
                                           name != 'Altona Meadows') %>% st_union()
  
  sub_sdf <- ltr_tdf %>%
    group_by(suburb.x, Submarket) %>%
    dplyr::summarize(count=n())
  sub_sf <- suburb_sf %>%
    dplyr::left_join(sub_sdf %>%
                       dplyr::select(name = suburb.x, Submarket))
  
  cc_sf <- sub_sf %>% dplyr::filter(Submarket == '   City-Core   ') %>% st_union()
  beach_sf <- sub_sf %>% dplyr::filter(Submarket == '   Beach   ') %>% st_union()
  
   # Make map
  
  png(filename = file.path(getwd(), 'images', "sub_loc_map.png"), 
       width = 4.5, height = 4.5, units = 'in', res = 300)

  tm_shape(suburb_sf %>% dplyr::filter(class != 'Outer')) +
    tm_polygons("class", palette=c('gray50', 'gray80'), title='Suburb Class') +
  tm_shape(studyarea_sf) + 
    tm_borders(col='black', lwd=5) +
    tm_add_legend(type='fill', border.lwd=3, col='white', border.col='black',
                  title='Study Area')+
  tm_layout(
    legend.position = c("left","bottom"),
    legend.text.size = 0.5)
  
  dev.off()  
  
```

## Point Location Maps

We then add the short and long term observation locations to the maps. 

```{r location map}


  ltr_map <-   tm_shape(mid_sf) + 
    tm_fill(col='gray95', lwd=1) +
    tm_borders(col = 'gray40', lwd=1) +
  tm_shape(inn_sf) + 
    tm_fill(col='gray85', lwd=1) +  
  tm_shape(cc_sf) + 
    tm_fill(col='gray75', lwd=1) +
  tm_shape(beach_buffer) + 
    tm_fill(col='gray65', lwd=1) +
  tm_shape(suburb_sf %>% dplyr::filter(in_studyarea == 1)) +
    tm_borders(col='gray70', lwd=.5) +
  tm_shape(studyarea_sf) + 
    tm_borders(col='black', lwd=2) +
  tm_shape(ltr_sf %>% dplyr::rename(Submarket = geo_mrkt)) +
    tm_symbols('Submarket', size=.02,
               palette = sm_col, alpha=.9, shape=16, legend.col.show=TRUE)+
    tm_layout(title = "LTR Observations",
              title.position = c("left", "top"),
              title.size = .8) +
  tm_shape(inn_sf) + 
    tm_borders(col = 'gray40', lwd=1) +
  tm_shape(cc_sf) + 
    tm_borders(col = 'gray40', lwd=1) +
  tm_shape(beach_buffer) + 
    tm_borders(col = 'gray40', lwd=1)
  
  
  str_map <-   tm_shape(mid_sf) + 
    tm_fill(col='gray95', lwd=1) +
    tm_borders(col = 'gray40', lwd=1) +
  tm_shape(inn_sf) + 
    tm_fill(col='gray85', lwd=1) +  
  tm_shape(cc_sf) + 
    tm_fill(col='gray75', lwd=1) +
  tm_shape(beach_buffer) + 
    tm_fill(col='gray65', lwd=1) +
  tm_shape(suburb_sf %>% dplyr::filter(in_studyarea == 1)) +
    tm_borders(col='gray70', lwd=.5) +
  tm_shape(studyarea_sf) + 
    tm_borders(col='black', lwd=2) +
  tm_shape(str_sf %>% dplyr::rename(Submarket = geo_mrkt)) +
    tm_symbols('Submarket', size=.02,
               palette = sm_col, alpha=.9, shape=16, legend.col.show=TRUE)+
    tm_layout(title = "STR Observations",
              title.position = c("left", "top"),
              title.size = .8) +
  tm_shape(inn_sf) + 
    tm_borders(col = 'gray40', lwd=1) +
  tm_shape(cc_sf) + 
    tm_borders(col = 'gray40', lwd=1) +
  tm_shape(beach_buffer) + 
    tm_borders(col = 'gray40', lwd=1)
  
  
  sltr_map <- tmap_arrange(ltr_map, str_map) 
    
  png(filename = file.path(getwd(), 'images', "ltr_loc_map.png"), 
       width = 7, height = 3.5, units = 'in', res = 300)
    sltr_map
  dev.off()  
    


```



### Revenue Density

Here we make the density plots of the short and long term observed revenues. 

```{r maps}

   dens_cols <- c(col_df$rgb[3], col_df$rgb[1], col_df$rgb[2])

   # Add together for plotting
   rev_df <- bind_rows(str_tdf %>% 
                         dplyr::select(type, revenue=revenue_net) %>%
                         dplyr::mutate(tenure = '  Short Term (Actual)   ',
                                       alpha = .5),
                       str_tdf %>% 
                         dplyr::select(type, revenue=pot_revenue_net) %>%
                         dplyr::mutate(tenure = '  Short Term (Potential)   ',
                                       alpha = .4),
                       ltr_df %>%
                         dplyr::select(type, revenue=gross_revenue) %>%
                         dplyr::mutate(tenure = '  Long-Term   ',
                                       alpha = .9))

  png(filename = file.path(getwd(), 'images', "rev_dens.png"), 
       width = 5.5, height = 4, units = 'in', res = 300)
 
   ggplot(rev_df,
          aes(x=revenue, fill=tenure, color=tenure)) +
    geom_density(alpha=.5) +
    scale_fill_manual(values=dens_cols) +
    scale_color_manual(values=dens_cols) +
    # ggtitle('Revenue Comparison: Short Term vs Long Term Properties') +
    xlab('Annual Revenue (A$)') +
    ylab('')+
    scale_x_continuous(breaks=c(seq(0,60000, by=20000)),
                       labels=c('A$0', 'A$20,000', 'A$40,000', 'A$60,000'))+
    theme(legend.position='bottom',
          legend.title = element_blank())+
     theme(plot.title = element_text(size = 10, hjust = 0.5),
           axis.text.y=element_blank(),
           axis.text.x=element_text(size = 6),
           axis.ticks.y=element_blank(),
           plot.caption=element_text(size=6, hjust=0)) + 
     labs(caption = paste0('\nComparison of the distribution of observed long-term rental',
                           ' revenues of 64,305 properties (blue) to observed revenues of ', 
                           ' 2,415 short\nterm properties (pink).  The purple curve ',
                           ' represents the potential revenues of the 2,415 short term ',
                           ' properties (includes imputed bookings).')) +
    coord_cartesian(xlim=c(0, 70000)) 

    dev.off()  
        
```

### Compare Imputation methods

We build a plot to compare the two imputation methods, hedonic and KNN. 

```{r message=FALSE, warning=FALSE, comment=FALSE, cache=FALSE, echo=FALSE}
    
  library(ggplot2)

   load(file.path('c:/code', 'research', 'airbnbMelbourne',
                  'data', 'analyzed', 'reviewers.RData'))

  imp_df$bed_lab <- paste0(imp_df$bedrooms, ' Beds')
  fig_cap <- paste0('This figure compares the estimated long term rent for short term rental properties via the hedonic imputation\nmethod (X-axis) and the matched pairs-type method (Y-axis).  It is broken down by bedroom count and\nstructure type (Apartment vs House). Overall correlation is 89.4%')
  
  png(filename = file.path(getwd(), 'images', "imp_cor.png"), 
       width = 7, height = 4, units = 'in', res = 300)

  ggplot(imp_df,
         aes(x=ltr_imp_revenueH, y=ltr_imp_revenueK, col = type)) + 
    geom_point(alpha = .5, size=.5) + 
    facet_grid(type ~ bed_lab) + 
    xlab('\nHedonic Revenue (A$)\n') + 
    ylab('Matched Pairs Revenue (A$)\n') + 
    theme(legend.position = 'none',
          plot.caption=element_text(hjust = 0),
          plot.title = element_text(hjust = 0.5)) +
    scale_color_manual(values = c(col_df$rgb[8], col_df$rgb[10])) + 
    scale_x_continuous(breaks = c(15000, 30000, 45000),
                       labels = c('$15k', '$30k', '$45k'))+
    scale_y_continuous(breaks = c(15000, 30000, 45000),
                       labels = c('$15k', '$30k', '$45k'))+
    geom_abline(intercept = 0, slope = 1, linetype = 2,
                color = 'gray50') +
    labs(caption = fig_cap)
 
  dev.off()  
   

```

## Conditional Expectation Plots

Finally, we build the conditional expectation plots.  

First, the 1 bed/1bath option. 

```{r}
 ## 1/1 plots
  x <- pred_df[pred_df$bedbath == "1_1", ]
  cap11 <- paste0('Plot shows the conditional probability of short term preference for ',
                  '1 bedroom, 1 bathroom homes given differences\n',
                  'in their structure type (Apt vs House), cancellation policy (strict',
                  ' vs flexible), location (City-Core, City, Beach, Suburban),\n',
                  'Number of allowed guest per bedroom and minimum number of night stay ',
                  'reguired. Larger and darker dots indicate\na higher probabilty ',
                  'of short-term preference with full dots showing 50%+.')

  ggplot(x, aes(x=ms, y=gpb, color = prob)) + 
    geom_point(aes(shape = sh, size = prob)) + 
    facet_grid(sm~ct) +
    scale_color_gradient(low='gray80', high='black', name='STP Probability',
                         labels = c('0%', '25%', '50%', '75%'),
                         breaks = c(0,.25, .50, .75)) +
    scale_shape_manual(values = c(16,1), name = 'Is STP?', labels=c('True', 'False')) + 
    scale_size(range = c(.5, 3), guide = FALSE) + 
    xlab('\n Minimum Stay  (In Days)\n') +
    ylab('Allowed Guests per Bedroom\n') + 
    labs(title = '1 Bed, 1 Bath Home - Conditional STP Probabilities\n',
         caption = cap11) + 
    theme(plot.caption=element_text(hjust = 0),
          plot.title = element_text(hjust = 0.5)) -> cpp_11
  
  
  png(filename = file.path(getwd(), 'images', "cpp_11.png"), 
       width = 7, height = 5, units = 'in', res = 300)
     cpp_11
  dev.off()
       
```

And then the 3bed, 2bath option.  

```{r}

 ## 3/2 plots
  x <- pred_df[pred_df$bedbath == "3_2", ]
  cap11 <- paste0('Plot shows the conditional probability of short term preference for ',
                  '3 bedroom, 2 bathroom homes given differences\n',
                  'in their structure type (Apt vs House), cancellation policy (strict',
                  ' vs flexible), location (City-Core, City, Beach, Suburban),\n',
                  'Number of allowed guest per bedroom and minimum number of night stay ',
                  'reguired. Larger and darker dots indicate\na higher probabilty ',
                  'of short-term preference with full dots showing 50%+.')

  ggplot(x, aes(x=ms, y=gpb, color = prob)) + 
    geom_point(aes(shape = sh, size = prob)) + 
    facet_grid(sm~ct) +
    scale_color_gradient(low='gray80', high='black', name='STP Probability',
                         labels = c('0%', '25%', '50%', '75%'),
                         breaks = c(0,.25, .50, .75)) +
    scale_shape_manual(values = c(16,1), name = 'Is STP?', labels=c('True', 'False')) + 
    scale_size(range = c(.5, 3), guide = FALSE) + 
    xlab('\n Minimum Stay (In Days)\n') +
    ylab('Allowed Guests per Bedroom\n') + 
    labs(title = '3 Bed, 2 Bath Home - Conditional STP Probabilities\n',
         caption = cap11) + 
    theme(plot.caption=element_text(hjust = 0),
          plot.title = element_text(hjust = 0.5)) -> cpp_32
  
  
  png(filename = file.path(getwd(), 'images', "cpp_32.png"), 
       width = 7, height = 5, units = 'in', res = 300)
     cpp_32
  dev.off()
       
```

